---
title: "p8105_hw3_pl2848"
author: "Pei Liu"
date: "2022-10-12"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
import all the library needed
```{r}
library(tidyverse)
library(p8105.datasets)
library(dplyr)
```

# Problem 1
```{r}
data("instacart")
```

1. How many aisles are there, and which aisles are the most items ordered from?
2. Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
3. Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
4. Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

# Problem 2

Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).
```{r}
accel_data = read_csv("data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(activity_1:activity_1440,
               names_to = "minutes",
               names_prefix = "activity_",
               values_to = "activity_counts") %>% 
  mutate(day = as.factor(day),
         day_type = 
           as.factor(ifelse(
             day == "Saturday" | day == "Sunday",
             "weekend", "weekday")),
         day_id = as.numeric(day_id),
         week = as.numeric(week),
         minutes = as.numeric(minutes))

nrow(accel_data)
ncol(accel_data)
head(accel_data, 10)

# Our tidy data set has 6 variables: week, day_id, day, minutes, activity_counts, and day_type.
# In total, there are 50400 observations.
# I also change the day and day_type to factorial variables, changed day_id, week and minutes to numeric variables.
```

Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?


```{r}

total_activity = accel_data %>% 
  group_by(day, week) %>% 
  summarise(total_counts = sum(activity_counts)) %>% 
  pivot_wider(
    names_from = day,
    values_from = total_counts
  ) 

knitr::kable(total_activity)
  


  
  
head(total_activity)


# I didn't find apparent trends. But the total activity on Saturday at week 4 and 5 is extrmely small compared to other days.
```



Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.
```{r}
total_activity_min = accel_data %>% 
  group_by(day, minutes) %>% 
  summarise(total_counts = sum(activity_counts))
total_activity_min

# I prepared the data for my ggplot in this step.
```


```{r}
trend_plot = total_activity_min %>% 
  ggplot(aes(x = minutes, y = total_counts, color = day)) + 
  geom_point(aes(color = day), alpha = .5, size = 0.1)  +
  scale_x_continuous(breaks = seq(0, 1440, by = 120),
                    labels = c("0:00", "2:00", "4:00", "6:00", "8:00", "10:00", "12:00", "14:00", "16:00", "18:00" , "20:00", "22:00", "24:00")
                     ) +
  labs(
    title = "Activity over the days",
    x = "Time",
    y = "Total activity counts",
    caption = "Data from Accelerometer") +
  theme(legend.position = "bottom")
  
trend_plot
trend_plot + geom_line()
# Looking at the scatter plot, we can see that the counts of activity is low midnight (during the sleep time), and starts to increase after about 6. Then the counts reach the first peak around 12:00 (11:00. Then the counts drops around 13:00, and fluctuated until 20:00 to 22:00, where the activity reached the second peak. Then the activity drops after around 23:00.
```



# Problem 3
This problem uses the NY NOAA data. DO NOT include this dataset in your local data directory; instead, load the data from the p8105.datasets package using:
```{r}
# Loaded the data
library(p8105.datasets)
data("ny_noaa")
```

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. Then, do or answer the following (commenting on the results of each):



```{r}
# For this step, I seperate the date variable into year, month and day. I adjust the units for prcp by deviding original prcp by 10, and do the same thing for tmax and tmin.
library(lubridate)
ny_noaa_clean = ny_noaa %>% 
  mutate(year = year(date),
         month = month(date), 
         day = day(date),
         prcp = prcp/10 , 
         tmax = as.numeric(tmax)/10,  
         tmin = as.numeric(tmin)/10) %>% 
  arrange(date)
  

head(ny_noaa_clean)
```

```{r}
# In this step, I calculated the frequency of snow (mm)
snow_freq = ny_noaa_clean %>% 
  drop_na(snow) %>% 
  count(snow) %>% 
  arrange(desc(n))

snow_freq

# For snowfall, what are the most commonly observed values? Why?
# The most commonly observed values are 0. Because it wouldn't snow everyday, especially in seasons other than winter. So most of the values are 0.
```
Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?
```{r}
# I prepare the data for ploting. I only drop tmax = na here, because that's the variable that I am interested in. I also changed month from numer to its name. 
ny_noaa_clean %>% 
  drop_na(tmax) %>% 
  filter(month == 1 | month == 7) %>% 
  mutate (month = month.name[month]) %>% 
  group_by(year, month, id) %>% 
  summarize(avg_tmax = mean(tmax)) %>% 
# x-axis is year, y axis is avg_tmax. I also separate them by month, so that there are two plots.
  ggplot(aes(x = year, y = avg_tmax)) +
  geom_point(aes(color = month), size = 0.5)+
  theme(legend.position = "none")+
  facet_grid(.~ month) + 
  labs(
    title = "Average max temperature in January and in July across years",
    x = "Year",
    y = "Max Temperature") +
  geom_smooth()

# The temperatures in January is significantly lower than it is in July. The mean temperature of January fluctuated around 0 Celsius, and the range of the temperature is wider. The temperature in July over years is more stable, with less fluctuation around 27 Celsius, with less range. 
# We can see from the plot that there are two obvious outliers in January around 1982 and 2005 The temperatures were below -10 Celsius. For temperatures in July, there is one obvious outlier around 1988, the temperature was below 15 Celsius. 
```

Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.
```{r}
library(ggridges)
# (i) tmax vs tmin for the full dataset
p1 = ny_noaa_clean %>%  ggplot(aes(x = tmax, y = tmin)) + 
  geom_hex()

# (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year
p2 = ny_noaa_clean %>%  
  filter(snow > 0 & snow < 100) %>% 
  select(year, snow) %>% 
  ggplot(aes(x = snow, y = as.factor(year))) +
  geom_boxplot() +
  labs(
    title = "Snowfall (mm) among year",
    x = "Snowfall (mm)",
    y = "Year")
  
# I loaded library patchwork here to create the two penel plot
library(patchwork)
p1+p2 
```

